Enhancing \Hinf{} Norm Estimation using Local LPM/LRM Modeling: Applied to an AVIS\thanksref{thx}

### 2016-06-27 ch04-lrmhinf/lrmhinf.tex
Small corrections Ch04

To estimate $\gamma$, only $\Delta(\omega_{\star})$ is of interest, where
\begin{equation}
 \omega_{\star} 
     \isdef 
       \arg
         \sup_{\omega} 
           \svdMax{\Delta(\omega)}
  \text{.}
  \label{eq:lrmhinf:omega-star}
\end{equation}
If the measurement time is finite, non-parametric approaches only give access to a finite and discrete grid of frequencies $\omega_k \in \DFTGrid $, i.e. $\DFTGrid \isdef \set{\omega_1, \omega_2, \ldots, \omega_{N-1}}$.
Typically, this discrete grid is the frequency grid of the \gls{DFT}.
As such, \eqref{eq:lrmhinf:HinfNorm} can be rewritten as
\begin{equation}
  \gamma \isdef 
    \sup_{\omega} 
      \svdMax{\Delta(\omega)}
         = \max 
             \Set{
                \max_{\omega \in \DFTGrid  } 
                  \svdMax{\Delta(\omega)}
                \;,\;
                \sup_{\omega \in \InterGrid} 
                  \svdMax{\Delta(\omega)}
                }
  \label{eq:lrmhinf:gamma-split-grid}
  \text{.}
\end{equation}
For convenience, we extract both contributions above as:
\begin{align}
  \gammaAtGrid & \isdef 
    \max_{\omega \in \DFTGrid  } 
      \svdMax{\Delta(\omega)}
  \label{eq:lrmhinf:gamma:atGrid}\\
  \gammaIntergrid & \isdef 
    \sup_{\omega \in \InterGrid} 
      \svdMax{\Delta(\omega)}
   \label{eq:lrmhinf:gamma:interGrid}
\end{align}
where $\Omega \isdef \OpenInterval{0,\frac{2\pi}{\Ts}}$ is the measured frequency range for a sampling time $\Ts$.
Direct inspection of the \gls{FRF} data only allows access to $\gammaAtGrid$.
Two cases can be distinguished:
\begin{enumerate}
  \item $\omega_{\star} \in \DFTGrid$ such that $\svdMax{\Delta(\omega_{\star})} = \gamma =\gamma_{\FRF}$ is observed directly on the discrete frequency grid,
  \item $\omega_{\star} \not\in \DFTGrid$, i.e. the extremal value $\svdMax{\Delta(\omega_{\star})}  = \gamma = \gammaIntergrid$ is determined by the inter-grid behavior.
\end{enumerate}

@@

\section{\Hinf{} Norm Estimation Procedure}
\label{sec:lrmhinf:LPMHinf}
To estimate $\gammaIntergrid$, we propose to identify local parametric models in the frequency domain.
The frequency axis is partitioned into continuous segments $\Omega_k \isdef \ClosedInterval{ \omega_{k-1} , \omega_{k+1} }$ in view of the local validity of the local models.
In particular, the inter-grid aspect of the $\Hinf$ norm is estimated as
\begin{equation}
  \gammaIntergrid = 
    \sup \Set{ 
                \sup_{\omega \in \Omega_k} \svdMax{\LocalModel[k(\omega)]{\Delta}}
                |
                k = 1,\ldots,N-1
                }
   \label{eq:lrmhinf:gamma-inter-grid}
\end{equation}
where the $\LocalModel[k]{\Delta}$ is the local model around the ($k^{\text{th}}$) frequency bin $\omega_k$.
An example of such local models is shown in \figref{fig:lrmhinf:interpol-quantities}.

@@

\paragraph{Interpolation method}
To allow inspection of $\svdMax{\Delta(\omega)}$ in~\eqref{eq:lrmhinf:HinfNormSISO} for arbitrary values of $\omega \in \Omega$, we propose an interpolation approach to estimate $\gamma$ and to allow visual inspection of $\svdMax{\Delta(\omega)}$.
It can be understood intuitively that due to the limited validity domain of the local models, an estimate of $\Delta(\omega)$ can only be obtained from $\LocalModel[k(\omega)]{\Delta}$ when the distance $\abs{\omega_k-\omega}$ is small.
As the local models are each constructed around the frequency $\omega_k$, for an arbitrary $\omega \in \Omega$, one can find two adjacent frequencies around which such local models are available.
To this end, two functions that indicate respectively the left and the right adjacent local models for frequency $\omega \in \Omega$ are defined:
\begin{align}
  \leftAdjacent(\omega) & 
                \isdef
                  \max
                  \Set{
                    k 
                    |
                      \omega_k \in \DFTGrid,  \omega_k \leq \omega
                  }\\
  \rightAdjacent(\omega) & 
                \isdef
                  \min
                  \Set{
                    k
                    |
                      \omega_k \in \DFTGrid,  \omega_k \geq \omega
                  }
  \text{.}
  \label{eq:lrmhinf:kL-kR-indexing-functions}
\end{align}
Using the $\leftAdjacent(\omega)$ and $\rightAdjacent(\omega)$ functions, two candidate values for $\svdMax{\Delta(\omega)}$ are obtained:
\begin{equation*}
 \svdMax{\LocalModel[\leftAdjacent(\omega)]{\Delta}(\omega)} 
 \quad \text{ and } \quad
 \svdMax{\LocalModel[\rightAdjacent(\omega)]{\Delta}(\omega)}
\end{equation*}
In view of obtaining the maximum singular value, these are `interpolated' to obtain a single estimate:
\begin{equation}
  \svdMax{\Delta_{\LMM}(\omega)}
    \isdef
    \max \Set{
       \svdMax{\LocalModel[\leftAdjacent(\omega)]{\Delta}(\omega)},
        \svdMax{\LocalModel[\rightAdjacent(\omega)]{\Delta}(\omega)}
    }
    \label{eq:lrmhinf:DeltaLRMInterpolation}
\end{equation}
by retaining the largest of both.
It follows that
\begin{align}
  \widehat{\infnorm{\Delta_{\LMM}}}
    & \isdef
    \sup_{\omega\in\Omega}
    \svdMax{\Delta_{\LMM}(\omega)} \\
   &  =
      \sup_{\omega\in\Omega} \; 
      \max \Set{
                 \svdMax{\LocalModel[\leftAdjacent(\omega)]{\Delta}(\omega)}, 
                 \svdMax{\LocalModel[\rightAdjacent(\omega)]{\Delta}(\omega)}}
  \label{eq:lrmhinf:gammaInterpolOptim}
  \text{.}
\end{align}

@@

\begin{figure}
  \centering
  \setlength{\figurewidth}{0.75\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \input{\thisDir/figs/example-both.tikz}
  \caption[Example: \glsentryshort{FRF} underestimates $\E{\gamma}$]{Simulation example revealing that $\gamma_{\FRF}$ (left) for both the \glsfirst{SA}~\legref{leg:example-SA} and \gls{LRM}~\legref{leg:example-LRM} underestimate the true $\gamma = \infnorm{\Delta} = 24$~\legref{leg:example-true-Hinf} due to limited frequency resolution.
  Using the proposed \gls{LRM}-based approach, both the at-grid~($\gamma_{\FRF}$) and intergrid~($\gammaIntergrid$) estimate are improved.
  The interpolated \gls{LRM}~\legref{leg:example-LRM-interpol} and true $\Delta$~\legref{leg:example-true} almost coincide.
  Hence, $\widehat{\infnorm{\Delta_{\LRM}}}_{\infty}$~\legref{leg:example-LRM-interpol-Hinf} approaches the theoretical $\gamma=24$.}
  \label{fig:lrmhinf:exampleIntergrid}
\end{figure}

@@

\section{Application to an \glsentryname{AVIS}}
\label{sec:lrmhinf:measurement}
The proposed method is illustrated on the \gls{AVIS} shown in \figref{fig:lrmhinf:avis}.
The broad aim of vibration isolation~\citep{Karnopp1995,Collette2011} is to suppress vibrations such that a certain payload---e.g. passengers in a car, sensitive equipment such as atomic force microscopes and wafer-scanners used for lithography---is not disturbed by vibrations generated by either the payload itself ($\systemDisturbance$) or by the outside world ($\systemDisturbance$) as indicated in \figref{fig:lrmhinf:vibration-isolation}.

\begin{figure}
 \centering
  \setlength{\figurewidth}{0.45\columnwidth}
  \begin{subfigure}[b]{0.49\columnwidth}
    \input{\thisDir/figs/vibrationisolation-classical.tikz}
    \caption{Traditional set-up.}  
    \label{fig:lrmhinf:vibration-isolation:traditional}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\columnwidth}
    \input{\thisDir/figs/vibrationisolation-skyhook.tikz}
    \caption{Principle of sky-hook damping.}  
    \label{fig:lrmhinf:vibration-isolation:skyhook}
  \end{subfigure}
  \disclaimer{Figure based on \citep[Fig. 3 and 4]{Voorhoeve2015SYSID} and \citep[Fig. 3 and 6]{Karnopp1995}.}
  \caption[Vibration isolation set-ups (traditional, sky-hook dampinng)]{Conceptual illustration of (active) vibration isolation set-ups. 
  In addition to the passive isolation (spring and damper), the active controller $C$ provides additional isolation using the measurement of the payload velocity ($y=v_z$).
  % If the velocity relative to ground is measured (a), the controller design involves trading-off which disturbance should be suppressed better.
  % When the absolute velocity the payload is measured (b), the controller can suppress both disturbances as it acts as a damper between a virtual fixed point (``sky'') and the payload.
  }
  \label{fig:lrmhinf:vibration-isolation}
\end{figure}

@@

The particular \gls{AVIS} used for the experiments, consists of a support fixed to the ground and a payload platform, suspended on airmounts (see \figref{fig:lrmhinf:avis:details}) providing passive vibration isolation.
The payload platform is able to move in six mechanical degrees of freedom.
Six geophones measure the payload velocity in an inertial reference frame, hence providing absolute velocity measurements, and eight linear motors (see \figref{fig:lrmhinf:avis:details}) allow to actively compensate for vibration of the platform using the principle of `sky-hook damping'.
Please refer to \appref{app:avis-setup} for more details regarding the mechanical set-up, placement of the sensor and the signal processing involved.
To facilitate the presentation, only the vertical translative velocity ($v_z$), i.e. a \gls{SISO} system, is considered.

@@

\subsection{Measurement \& Control Framework}
\label{sec:lrmhinf:control}
The experiments on the \gls{AVIS} are carried out in closed-loop (\figref{fig:lrmhinf:tpc}) with $r_2 = 0$.
The controller $\Controller=\experimental\Controller$ is a diagonal \gls{PI} regulator which stabilizes the \gls{AVIS} but yields only moderate vibration isolation.
See \appref{app:avis-setup} for detailed information regarding both the experimental setup and the experimental controller.
The $r_1$ input consists of five periods of a random phase multisine of $65\,536$ samples with excited bins such that $\omega_k \approx 1.001 \omega_{k-1}$  and $\Ts=1\unit{ms}$.
See \chapref{sec:excitation} and \citep{Geerardyn2013TIM} for more information regarding the design of such a signal.

@@

The technique is illustrated on a simulation example where the \gls{LRM}-based interpolation yields a n accurate estimate of the $\Hinf{}$ norm.
It is observed that the measurement time is reduced by almost an order of magnitude compared with techniques that only examine $\Delta(\omega)$ on the \gls{DFT} frequency grid.
Using measurements on an \gls{AVIS}, it is illustrated that the interpolation-based results can substantially improve $\infnorm{\Delta}$ and/or reduce the measurement time four-fold.
Furthermore, the results are validated by a detailed validation measurement.
Ongoing research focuses on extending the technique to \gls{MIMO} systems.

>>>>

To estimate $\gamma$, only $\Delta(\omega_{\star})$ is of interest, where
\begin{equation}
 \omega_{\star} 
     \isdef 
       \arg
         \sup_{\omega} 
           \svdMax{\Delta(\omega)}
  \text{.}
  \label{eq:lrmhinf:omega-star}
\end{equation}
If the measurement time is finite, non-parametric approaches only give access to a finite and discrete grid of frequencies $\omega_k \in \DFTGrid $, i.e. $\DFTGrid \isdef \set{\omega_1, \omega_2, \ldots, \omega_{N-1}}$.
Typically, this discrete grid is the frequency grid of the \gls{DFT}.
As such, \eqref{eq:lrmhinf:HinfNorm} can be rewritten as
\begin{equation}
  \gamma \isdef 
    \sup_{\omega} 
      \svdMax{\Delta(\omega)}
         = \max 
             \Set{
                \max_{\omega \in \DFTGrid  } 
                  \svdMax{\Delta(\omega)}
                \;,\;
                \sup_{\omega \in \InterGrid} 
                  \svdMax{\Delta(\omega)}
                }
  \label{eq:lrmhinf:gamma-split-grid}
  \text{.}
\end{equation}
For convenience, we extract both contributions above as:
\begin{align}
  \gammaAtGrid & \isdef 
    \max_{\omega \in \DFTGrid  } 
      \svdMax{\Delta(\omega)}
  \label{eq:lrmhinf:gamma:atGrid}\\
  \gammaIntergrid & \isdef 
    \sup_{\omega \in \InterGrid} 
      \svdMax{\Delta(\omega)}
   \label{eq:lrmhinf:gamma:interGrid}
\end{align}
where $\Omega \isdef \OpenInterval{0,\frac{2\pi}{\Ts}}$ is the measured frequency range for a sampling time $\Ts$.
Direct inspection of the \gls{FRF} data only allows access to $\gammaAtGrid$.
Two cases can be distinguished:
\begin{enumerate}
  \item $\omega_{\star} \in \DFTGrid$ such that $\svdMax{\Delta(\omega_{\star})} = \gamma =\gamma_{\FRF}$ is observed directly on the discrete frequency grid,
  \item $\omega_{\star} \not\in \DFTGrid$, i.e. the extremal value $\svdMax{\Delta(\omega_{\star})}  = \gamma = \gammaIntergrid$ is determined by the \gls{IG} behavior.
\end{enumerate}

@@

\section{\Hinf{} Norm Estimation Procedure}
\label{sec:lrmhinf:LPMHinf}
To estimate $\gammaIntergrid$, we propose to identify local parametric models in the frequency domain.
The frequency axis is partitioned into continuous segments $\Omega_k \isdef \ClosedInterval{ \omega_{k-1} , \omega_{k+1} }$ in view of the local validity of the local models.
In particular, the inter-grid aspect of the $\Hinf$ norm is estimated as
\begin{equation}
  \gammaIntergrid = 
    \sup \Set{ 
                \sup_{\omega \in \Omega_k} \svdMax{\LocalModel[k]{\Delta}(\omega)}
                |
                k = 1,\ldots,N-1
                }
   \label{eq:lrmhinf:gamma-inter-grid}
\end{equation}
where the $\LocalModel[k]{\Delta}$ is the local model around the ($k^{\text{th}}$) frequency bin $\omega_k$.
An example of such local models is shown in \figref{fig:lrmhinf:interpol-quantities}.

@@

\paragraph{Interpolation method}
To allow inspection of $\svdMax{\Delta(\omega)}$ in~\eqref{eq:lrmhinf:HinfNormSISO} for arbitrary values of $\omega \in \Omega$, we propose an interpolation approach to estimate $\gamma$ and to allow visual inspection of $\svdMax{\Delta(\omega)}$.
It can be understood intuitively that due to the limited validity domain of the local models, an estimate of $\Delta(\omega)$ can only be obtained from $\LocalModel[k(\omega)]{\Delta}$ when the distance $\abs{\omega_k-\omega}$ is small.
As the local models are each constructed around the frequency $\omega_k$, for an arbitrary $\omega \in \Omega$, one can find two adjacent frequencies around which such local models are available.
To this end, two functions that indicate respectively the left and the right adjacent local models for frequency $\omega \in \Omega$ are defined:
\begin{align}
  \leftAdjacent(\omega) & 
                \isdef
                  \max
                  \Set{
                    k 
                    |
                      \omega_k \in \DFTGrid,  \omega_k \leq \omega
                  }\\
  \rightAdjacent(\omega) & 
                \isdef
                  \min
                  \Set{
                    k
                    |
                      \omega_k \in \DFTGrid,  \omega_k \geq \omega
                  }
  \text{.}
  \label{eq:lrmhinf:kL-kR-indexing-functions}
\end{align}
Using the $\leftAdjacent(\omega)$ and $\rightAdjacent(\omega)$ functions, two candidate values for $\svdMax{\Delta(\omega)}$ are obtained:
\begin{equation*}
 \svdMax{\LocalModel[\leftAdjacent(\omega)]{\Delta}(\omega)} 
 \quad \text{ and } \quad
 \svdMax{\LocalModel[\rightAdjacent(\omega)]{\Delta}(\omega)}
\end{equation*}
In view of obtaining the maximum singular value, these are `interpolated' to obtain a single estimate:
\begin{equation}
  \svdMax{\Delta_{\LMM}(\omega)}
    \isdef
    \max \Set{
       \svdMax{\LocalModel[\leftAdjacent(\omega)]{\Delta}(\omega)},
        \svdMax{\LocalModel[\rightAdjacent(\omega)]{\Delta}(\omega)}
    }
    \label{eq:lrmhinf:DeltaLRMInterpolation}
\end{equation}
by retaining the largest of both.
%TODO: remark that this notation is deceiving: there is no $\Delta_{\LMM}$, and hence, its Hinf norm does not exist. Both notations are just to indicate
It follows that
\begin{align}
  \widehat{\infnorm{\Delta_{\LMM}}}
    & \isdef
    \sup_{\omega\in\Omega}
    \svdMax{\Delta_{\LMM}(\omega)} \\
   &  =
      \sup_{\omega\in\Omega} \; 
      \max \Set{
                 \svdMax{\LocalModel[\leftAdjacent(\omega)]{\Delta}(\omega)}, 
                 \svdMax{\LocalModel[\rightAdjacent(\omega)]{\Delta}(\omega)}}
  \label{eq:lrmhinf:gammaInterpolOptim}
  \text{.}
\end{align}

@@

\begin{figure}
  \centering
  \setlength{\figurewidth}{0.75\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \input{\thisDir/figs/example-both.tikz}
  \caption[Example: \glsentryshort{FRF} underestimates $\E{\gamma}$]{Simulation example revealing that $\gamma_{\FRF}$ (left) for both the \glsfirst{SA}~\legref{leg:example-SA} and \gls{LRM}~\legref{leg:example-LRM} underestimate the true $\gamma = \infnorm{\Delta} = 24$~\legref{leg:example-true-Hinf} due to limited frequency resolution.
  Using the proposed \gls{LRM}-based approach, both the at-grid~($\gamma_{\FRF}$) and intergrid~($\gammaIntergrid$) estimate are improved.
  The interpolated \gls{LRM}~\legref{leg:example-LRM-interpol} and true $\Delta$~\legref{leg:example-true} almost coincide.
  Hence, $\widehat{\infnorm{\Delta_{\LRM}}}$~\legref{leg:example-LRM-interpol-Hinf} approaches the theoretical $\gamma=24$.}
  \label{fig:lrmhinf:exampleIntergrid}
\end{figure}

@@

\section{Application to an \glsentryname{AVIS}}
\label{sec:lrmhinf:measurement}
The proposed method is illustrated on the \gls{AVIS} shown in \figref{fig:lrmhinf:avis}.
The broad aim of vibration isolation~\citep{Karnopp1995,Collette2011} is to suppress vibrations such that a certain payload---e.g. passengers in a car, sensitive equipment such as atomic force microscopes and wafer-scanners used for lithography---is not disturbed by vibrations generated by either the payload itself ($\systemDisturbance$) or by the outside world ($\externalDisturbance$) as indicated in \figref{fig:lrmhinf:vibration-isolation}.

\begin{figure}
 \centering
  \setlength{\figurewidth}{0.45\columnwidth}
  \begin{subfigure}[b]{0.49\columnwidth}
    \input{\thisDir/figs/vibrationisolation-classical.tikz}
    \caption{Traditional set-up.}  
    \label{fig:lrmhinf:vibration-isolation:traditional}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\columnwidth}
    \input{\thisDir/figs/vibrationisolation-skyhook.tikz}
    \caption{Principle of sky-hook damping.}  
    \label{fig:lrmhinf:vibration-isolation:skyhook}
  \end{subfigure}
  \disclaimer{Figure based on \citep[Fig. 3 and 4]{Voorhoeve2015SYSID} and \citep[Fig. 3 and 6]{Karnopp1995}.}
  \caption[Vibration isolation set-ups (traditional, sky-hook damping)]{Conceptual illustration of (active) vibration isolation set-ups. 
  In addition to the passive isolation (spring and damper), the active controller $C$ provides additional isolation using the measurement of the payload velocity ($y=v_z$).
  % If the velocity relative to ground is measured (a), the controller design involves trading-off which disturbance should be suppressed better.
  % When the absolute velocity the payload is measured (b), the controller can suppress both disturbances as it acts as a damper between a virtual fixed point (``sky'') and the payload.
  }
  \label{fig:lrmhinf:vibration-isolation}
\end{figure}

@@

The particular \gls{AVIS} used for the experiments, consists of a support fixed to the ground and a payload platform, suspended on airmounts (see \figref{fig:lrmhinf:avis:details}) providing passive vibration isolation.
The payload platform is able to move in six mechanical degrees of freedom.
Six geophones measure the payload velocity in an inertial reference frame, hence providing absolute velocity measurements, and eight linear motors (see \figref{fig:lrmhinf:avis:details}) allow to actively compensate for vibration of the platform using the principle of `sky-hook damping'.
Please refer to \appref{app:avis-setup} for more details regarding the mechanical set-up, placement of the sensors and the signal processing involved.
To facilitate the presentation, only the vertical translative velocity ($v_z$), i.e. a \gls{SISO} system, is considered.

@@

\subsection{Measurement \& Control Framework}
\label{sec:lrmhinf:control}
The experiments on the \gls{AVIS} are carried out in closed-loop (\figref{fig:lrmhinf:tpc}) with $r_2 = 0$.
%TODO: define T(P,C) formally
The controller $\Controller=\experimental\Controller$ is a diagonal \gls{PI} regulator which stabilizes the \gls{AVIS} but yields only moderate vibration isolation.
See \appref{app:avis-setup} for detailed information regarding both the experimental setup and the experimental controller.
The $r_1$ input consists of five periods of a random phase multisine of $65\,536$ samples with excited bins such that $\omega_k \approx 1.001 \omega_{k-1}$  and $\Ts=1\unit{ms}$.
See \chapref{sec:excitation} and \citep{Geerardyn2013TIM} for more information regarding the design of such a signal.

@@

The technique is illustrated on a simulation example where the \gls{LRM}-based interpolation yields an accurate estimate of the $\Hinf{}$ norm.
It is observed that the measurement time is reduced by almost an order of magnitude compared with techniques that only examine $\Delta(\omega)$ on the \gls{DFT} frequency grid.
Using measurements on an \gls{AVIS}, it is illustrated that the interpolation-based results can substantially improve $\infnorm{\Delta}$ and/or reduce the measurement time four-fold.
Furthermore, the results are validated by a detailed validation measurement.
Ongoing research focuses on extending the technique to \gls{MIMO} systems.
